Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-06-08 paper/paper.tex
Save a bit of space

\Cref{figure:connected-cp} compares the three approaches for ensuring connectedness in CP: by
branching (Branch), by filtering (Filter), or by combining branching and filtering (Both). We show
only results using the best variant for each class---that is, MAC for labelled graphs, and FC for
unlabelled graphs (the other results are very similar).  On labelled graphs, we see many instances
which are solved very quickly by branching but not at all by filtering, and vice versa.  However,
combining both is rarely much worse than just doing one or the other, and is often much better, even
if on average it is slightly slower.

On unlabelled graphs, we observe that the three variants have rather similar performance.

@@

\begin{algorithm}[tb] \DontPrintSemicolon
\nl $\FuncSty{associationMCCIS}$ :: (Graph $G_1$, Graph $G_2$) $\rightarrow$ Map \;
\nl \Begin{
    \nl $\KwSty{global} ~ \mathit{incumbent} \gets \emptyset$ \label{incumbent} \;
    \nl $G \gets G_1 \modularproduct G_2$ \label{buildassoc} \;
    \nl $\FuncSty{search}(G, \emptyset, \emptyset, \operatorname{V}(G))$ \label{firstsearch} \;
    \nl $\KwSty{return} ~ \mathit{incumbent}$ \;
}
\vspace{1ex}
\nl $\FuncSty{search}$ :: (Graph $G$, Set $\mathit{solution}$, Set $\mathit{connected}$, Set $\mathit{remaining}$) \;
\nl \Begin{
    \nl $\mathit{colourClasses} \gets \FuncSty{concatenate}($ \label{makecolours} \;
    $\hspace*{2em}\FuncSty{colour}(G, \mathit{remaining} \setminus \mathit{connected}),$ \;
    $\hspace*{2em}\FuncSty{colour}(G, \mathit{remaining} \cap \mathit{connected}))$ \;
    \nl \While{$\FuncSty{length}(\mathit{colourClasses}) > 0$\label{outerloop}}{
        \nl \ForEach{$v \in \FuncSty{last}(\mathit{colourClasses}) \textnormal{~in reverse order}$\label{innerloop}}{
            \nl \lIf{$|\mathit{solution}| + \FuncSty{length}(\mathit{colourClasses}) \leq
            |\mathit{incumbent}|$}{\KwSty{return}\label{bound}}
            \nl \lIf{$v \notin \mathit{connected} ~ \KwSty{and} ~ \mathit{solution} \ne
            \emptyset$}{\KwSty{return} \label{acceptbranchable}}
            \nl $\mathit{solution}' \gets \mathit{solution} \cup \{ v \} \label{addv} \label{addvtosolution} $\;
            \nl \lIf{$|\mathit{solution}'| > |\mathit{incumbent}|$}{$\mathit{incumbent} \gets
            \mathit{solution}'$ \label{newincumbent}}
            \nl $\mathit{connected}' \gets \mathit{connected} \cup \{ w \in G : \FuncSty{first}(w)
            \in \operatorname{N}(G, \FuncSty{first}(v)) \}$ \label{addtobranchable} \;
            \nl $\mathit{remaining}' \gets \mathit{remaining} \cap \operatorname{N}(G, v)$ \label{filterremaining} \;
            \nl \lIf{$\mathit{remaining}' \neq \emptyset$}{$\FuncSty{search}(G, \mathit{solution}',
            \mathit{connected}', \mathit{remaining}')$\label{recursesearch}}
         }
         \nl $\FuncSty{removeLast}(\mathit{colourClasses})$ \;
     }
}
\vspace{0.5ex}
\nl $\FuncSty{colour}$ :: (Graph $G$, Set $\mathit{uncoloured}$) $\rightarrow$ List of List of Vertex \;
\nl \Begin{
    \nl $\KwSty{return}$ a greedy sequential colouring of the vertices in $\mathit{uncoloured}$,
    using the procedure of San Segundo et al.\ \cite{DBLP:journals/cor/SegundoRJ11} with a static
    degree order from $G$ and $k_{\mathit{min}} = 0$. \;
}
\caption{An algorithm for a maximum common connected induced subgraph isomorphism via an association graph.}
\label{algorithm:mccis}
\end{algorithm}

>>>>

\Cref{figure:connected-cp} compares the three approaches for ensuring connectedness in CP: by
branching (Branch), by filtering (Filter), or by combining branching and filtering (Both). We show
only results using the best variant for each class---that is, MAC for labelled graphs, and FC for
unlabelled graphs (the other results are very similar).  On labelled graphs, we see many instances
which are solved very quickly by branching but not at all by filtering, and vice versa.  However,
combining both is rarely much worse than just doing one or the other, and is often much better, even
if on average it is slightly slower.  On unlabelled graphs, the three variants have rather similar
performance.

@@

\begin{algorithm}[tb] \DontPrintSemicolon
\nl $\FuncSty{associationMCCIS}$ :: (Graph $G_1$, Graph $G_2$) $\rightarrow$ Map \;
\nl \Begin{
    \nl $\KwSty{global} ~ \mathit{incumbent} \gets \emptyset$ \label{incumbent} \;
    \nl $G \gets G_1 \modularproduct G_2$ \label{buildassoc} \;
    \nl $\FuncSty{search}(G, \emptyset, \emptyset, \operatorname{V}(G))$ \label{firstsearch} \;
    \nl $\KwSty{return} ~ \mathit{incumbent}$ \;
}
\vspace{1ex}
\nl $\FuncSty{search}$ :: (Graph $G$, Set $\mathit{solution}$, Set $\mathit{connected}$, Set $\mathit{remaining}$) \;
\nl \Begin{
    \nl $\mathit{colourClasses} \gets \FuncSty{concatenate}($ \label{makecolours} \;
    $\hspace*{2em}\FuncSty{colour}(G, \mathit{remaining} \setminus \mathit{connected}),\,\FuncSty{colour}(G, \mathit{remaining} \cap \mathit{connected}))$ \;
    \nl \While{$\FuncSty{length}(\mathit{colourClasses}) > 0$\label{outerloop}}{
        \nl \ForEach{$v \in \FuncSty{last}(\mathit{colourClasses}) \textnormal{~in reverse order}$\label{innerloop}}{
            \nl \lIf{$|\mathit{solution}| + \FuncSty{length}(\mathit{colourClasses}) \leq
            |\mathit{incumbent}|$}{\KwSty{return}\label{bound}}
            \nl \lIf{$v \notin \mathit{connected} ~ \KwSty{and} ~ \mathit{solution} \ne
            \emptyset$}{\KwSty{return} \label{acceptbranchable}}
            \nl $\mathit{solution}' \gets \mathit{solution} \cup \{ v \} \label{addv} \label{addvtosolution} $\;
            \nl \lIf{$|\mathit{solution}'| > |\mathit{incumbent}|$}{$\mathit{incumbent} \gets
            \mathit{solution}'$ \label{newincumbent}}
            \nl $\mathit{connected}' \gets \mathit{connected} \cup \{ w \in G : \FuncSty{first}(w)
            \in \operatorname{N}(G, \FuncSty{first}(v)) \}$ \label{addtobranchable} \;
            \nl $\mathit{remaining}' \gets \mathit{remaining} \cap \operatorname{N}(G, v)$ \label{filterremaining} \;
            \nl \lIf{$\mathit{remaining}' \neq \emptyset$}{$\FuncSty{search}(G, \mathit{solution}',
            \mathit{connected}', \mathit{remaining}')$\label{recursesearch}}
         }
         \nl $\FuncSty{removeLast}(\mathit{colourClasses})$ \;
     }
}
\vspace{0.5ex}
\nl $\FuncSty{colour}$ :: (Graph $G$, Set $\mathit{uncoloured}$) $\rightarrow$ List of List of Vertex \;
\nl \Begin{
    \nl $\KwSty{return}$ a greedy colouring of the vertices in $\mathit{uncoloured}$,
    using the procedure of San Segundo et al.\ \cite{DBLP:journals/cor/SegundoRJ11} with a static
    degree order from $G$ and $k_{\mathit{min}} = 0$. \;
}
\caption{An algorithm for a maximum common connected induced subgraph isomorphism via an association graph.}
\label{algorithm:mccis}
\end{algorithm}
